---
title: "R Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
### Body Fat data
### Extra sum of squares
###
rm(list=ls())
library(ALSM)
```

```{r}
n <- dim(BodyFat)[1]
y  <- BodyFat$y
x1 <- BodyFat$x1
x2 <- BodyFat$x2
x3 <- BodyFat$x3
```

```{r}
## y ~x1
slr.x1 <- lm(y ~ x1)
coefficients(slr.x1)
anova(slr.x1)
```

```{r}
## y ~x2
slr.x2 <- lm(y ~ x2)
coefficients(slr.x2)
anova(slr.x2)
```

```{r}
## y ~x1+x2
slr.x1x2 <- lm(y ~ x1+x2)
coefficients(slr.x1x2)
anova(slr.x1x2)
```

```{r}
## y ~x1+x2+x3
slr.x1x2x3 <- lm(y ~ x1+x2+x3)
coefficients(slr.x1x2x3)
anova(slr.x1x2x3)
```

```{r}
### Allen test example
allen <- read.table("allen.txt",header = T)
names(allen)
```

```{r}
allen.lm123 <- lm(ACL ~ Vocab+ Abstract + SDMT , data=allen)
anova(allen.lm123)
```

```{r}
allen.lm23 <- lm(ACL ~ Abstract + SDMT , data=allen)
anova(allen.lm23)
```

```{r}
anova(allen.lm23,allen.lm123)
```

```{r}
allen.lm <- lm(ACL ~ Vocab, data=allen)# Vocab is x1
summary(allen.lm)
anova(allen.lm) # SSR? SSE? SStotal?
```

SSR(x1)=2.691 \#SSE(x1)=40.359 \#SST=43.050

```{r}
allen.lm13 <- lm(ACL ~ Vocab + SDMT, data=allen)      
summary(allen.lm13) 
anova(allen.lm13)
```

SSE(x1,x3)=31.2717 SSR(x1,x3)=2.6906+9.0872=11.7778, how does this
compare with previous SSR? SST=43.0495

stay the same. Why? it has nothing to do with predictors. Check the
formulas.

SSR(x3\|x1)= SSR(x1,x3)-SSR(x1)=11.7778-2.6906=9.0872 SSR(x3\|x1)=
SSE(x1)-SSE(x1,x3)=40.359-31.2717=9.0872

```{r}
anova(update(allen.lm13, . ~ . - SDMT),allen.lm13)
anova(update(allen.lm13, . ~ . - Vocab),allen.lm13)
```

```{r}
drop1(allen.lm13,test = "F")
```

does the order matter?

```{r}
allen.lm <- lm(ACL ~ SDMT, data=allen)
summary(allen.lm)
anova(allen.lm) 

```

SSR(x3)=11.68
SSE(x3)=31.37
SST=sum(anova(allen.lm)$'Sum Sq')=43.04957

SSR(x1|x3)=SSR(x1,x3)-SSR(x3)=11.7778-11.68=0.098
          =SSE(x3)-SSE(x1,x3)=31.37-31.2717=0.098
SSR(x1)=2.691 -->
 x1= Vocab doesn't contribute much to explain/reduce the variability in y 
after x3 = SDMT has already been considered. 

```{r}
names(allen)
```

```{r}
lm1<- lm(ACL ~ Vocab + SDMT + Abstract, data=allen)
anova(lm1)
```

```{r}
lm2<- lm(ACL ~ SDMT + Vocab + Abstract, data=allen)
anova(lm2)
```

```{r}
lm3<- lm(ACL ~  Abstract + SDMT + Vocab, data=allen)
anova(lm3)
```

```{r}
allen.lm123 <- lm(ACL ~ Vocab+ Abstract + SDMT , data=allen)
anova(allen.lm123)
```

```{r}
model0 <- lm(ACL ~ 1, data=allen)
anova(model0)
```

```{r}
anova(model0, allen.lm123)
```

```{r}
anova(update(allen.lm123, . ~ . - Vocab - Abstract - SDMT),allen.lm123)
anova(update(allen.lm123, . ~ . - Vocab),allen.lm123)
anova(update(allen.lm123, . ~ . - Abstract - SDMT),allen.lm123)
```

```{r}
anova(allen.lm, allen.lm123)
```

## rabbit example

```{r}

rabbit <- read.table("rabbit.txt",header = T)
names(rabbit)
colnames(rabbit)[1] <- "Y"
```

```{r}
rabbit.m1 <- lm(Y ~ Area + X2 + X3, data = rabbit)
summary(rabbit.m1)
anova(rabbit.m1)
```

```{r}
rabbit.m0 <-lm(Y ~ 1, data = rabbit)
anova(rabbit.m0,rabbit.m1)
```

```{r}
rabbit.m2 <- lm(Y  ~ Area, data = rabbit)
anova(rabbit.m2)
```

```{r}

((0.87926-0.54491)/(30-28)) / (0.54491/28)
```

```{r}
anova(rabbit.m2, rabbit.m1)
```

$$SSR(X2,X3|Area)$$

```{r}
0.31453+0.01981 # 
```

```{r}
0.33434/0.87926 
```
 Partial R-squared $$(X2,X3|Area)$$
              $$SSE(Area) = 0.87926$$
              

```{r}
rabbit.m3 <- lm(Y ~ X2+X3, data = rabbit)
anova(rabbit.m3)
anova(rabbit.m3,rabbit.m1)
```
